Title here

### 2020-09-01 analysis/paper/paper.tex
casi draft

We used monthly geopotential height at 2.5 longitude by 2.5 latitude
resolution from ERA5 \citep{hersbach} for the period 1979 to 2018
(inclusive).

@@

Figure \ref{fig:2d-regr} shows the spatial year-long regression \#FIXME.
Column a are regressinos using the Full SAM, while columns b and c are
regression coefficients computed in a multiple regression of
geopotential height on the Asymmetric and Symmetric indices at the same
time. Thus, they are to be interpreted as the patterns associated with
each index, controling for the (linear) effect of the other. (Figure A6
\#FIXME illustrates the difference between computing two simple
regressions and one multiple regression.)

In the stratosphere, the spatial pattern associated with the Full SAM is
more clearly dominated by a zonally symmetric, monopolar structure
(panel a.1) which is, however, not perfectly centered in the south pole.
The monopole obtained by multiple regression with the Asymmetric and
Symmetric SAM (panel c.1) is much more symmetric and the shift from
total symmetry is captured by the regression pattern of the Asymmetric
SAM as a wave-1 with maximum anomalies above the Belinghausen Sea on the
Western Hemisphere and and Davids Sea in the Eastern Hemisphere (panel
b.1).

In the troposphere, panel a.2 show the well known zonally symmetrical
annular mode \emph{contaminated} with zonal asymmetries in the form of a
wave-3. The regression using the Asymmetric and Symmetric SAM indices
successfully disentangle both structures. The Asymmetric component gives
rise to a cleaner zonal wave (panel b.2) and the Symemtric component is
assosiated with an trully annular mode, almost devoid of zonal
asymmetries (panel c.2). \#FIXME (agregar algo más?)

\begin{figure}
\includegraphics{wave-amplitude-1} \caption[Planteray wave amplitude for the regression patterns at 50 and 700 hPa]{Planteray wave amplitude for the regression patterns at 50 and 700 hPa.}\label{fig:wave-amplitude}
\end{figure}

The amplitude of each zonal wave number at each latitude at 50 hPa and
700 hPa is shown in Figure \ref{fig:wave-amplitude}, where wave number
zero represents the amplitude of the zonal mean. Comparing between rows,
this Figure quantifies the relatively clean separation between the
zonally symmetric and zonally asymmetric structures, as its evident how
the mixture of waves of the Full field (first row) is very similar to
the sum of the waves of the Asymmetric and Symmetric field (second and
third row, respectively). The second row of Figure
\ref{fig:wave-amplitude} shows that the Asymmetric SAM is overwhelmingly
dominated by wave 1 in the stratosphere (panel b), while in the
troposphere it is composed of zonal waves 3 to 1 in decreasing level of
importance.

@@

@@

\begin{sidewaysfigure}
\includegraphics{regr-air-season-1} \caption[Regression pattern of surface temperature with Asymmetric and Symmetric SAM]{Regression pattern of surface temperature with Asymmetric and Symmetric SAM. P-values smaller than 0.05 (controlling for Flase Detection Rate) as hatched areas. Gray areas have more than 15\% of missing data.}\label{fig:regr-air-season}
\end{sidewaysfigure}

@@

??

\begin{figure*}
\includegraphics{pp-regr-global-africa-1} \caption[Regression pattern of precipiation with Asymmetric and Symmetric SAM]{Regression pattern of precipiation with Asymmetric and Symmetric SAM. P-values smaller than 0.05 (controlling for Flase Detection Rate) as hatched areas. Africa}\label{fig:pp-regr-global-africa}
\end{figure*}

\begin{figure*}
\includegraphics{pp-regr-global-oceania-1} \caption[Same but for oceania]{Same but for oceania}\label{fig:pp-regr-global-oceania}
\end{figure*}

\begin{figure*}
\includegraphics{pp-regr-global-america-1} \caption[Same but for america]{Same but for america}\label{fig:pp-regr-global-america}
\end{figure*}

\subsection{Conclusiones}

@@

>>>>

We used monthly geopotential height at 2.5 longitude by 2.5 latitude
resolution from ERA5 \citep{hersbach} for the period 1979 to 2018.

@@

Figure \ref{fig:2d-regr} shows the spatial year-long regression \#FIXME.
Column a are regressinos using the Full SAM, while columns 2 and 3 are
regression coefficients computed in a multiple regression of
geopotential height on the Asymmetric and Symmetric indices at the same
time. Thus, they are to be interpreted as the patterns associated with
each index, controling for the (linear) effect of the other. (Figure A6
\#FIXME illustrates the difference between computing two simple
regressions and one multiple regression.)

In the stratosphere, the spatial pattern associated with the Full SAM is
more clearly dominated by a zonally symmetric, monopolar structure
(panel a.1) which is, however, not perfectly centered in the south pole.
The monopole obtained by multiple regression with the Asymmetric and
Symmetric SAM (panel a.3) is much more symmetric and the shift from
total symmetry is captured by the regression pattern of the Asymmetric
SAM as a wave-1 with maximum anomalies above the Belinghausen Sea on the
Western Hemisphere and and Davids Sea in the Eastern Hemisphere (panel
a.2).

In the troposphere, panel b.1 shows the well known zonally symmetrical
annular mode \emph{contaminated} with zonal asymmetries in the form of a
wave-3. The regression using the Asymmetric and Symmetric SAM indices
successfully disentangle both structures. The Asymmetric component gives
rise to a cleaner zonal wave (panel b.2) and the Symemtric component is
assosiated with an trully annular mode, almost devoid of zonal
asymmetries (panel b.3). note that the wave-3 pattern observed in panel
b.2 is rotated by half a wavelength from the average position of the
mean wave-3 pattern and asociated with \citet{raphael2004}'s ZW3 index
(see Figure 1 from that paper). \#FIXME (agregar algo más?)

\begin{figure}
\includegraphics{wave-amplitude-1} \caption[Planteray wave amplitude for the regression patterns at 50 and 700 hPa]{Planteray wave amplitude for the regression patterns at 50 and 700 hPa. Note the varying x axis.}\label{fig:wave-amplitude}
\end{figure}

The amplitude of each zonal wave number at each latitude at 50 hPa and
700 hPa is shown in Figure \ref{fig:wave-amplitude}, where wave number
zero represents the amplitude of the zonal mean. Comparing between rows,
this Figure quantifies the relatively clean separation between the
zonally symmetric and zonally asymmetric structures, as its evident how
the mixture of waves of the Full field (column a) is very similar to the
sum of the waves of the Asymmetric and Symmetric field (columns b and c,
respectively). Column b of Figure \ref{fig:wave-amplitude} shows that
the Asymmetric SAM is overwhelmingly dominated by wave 1 in the
stratosphere (panel b.1), while in the troposphere it is composed of
zonal waves 3 to 1 in decreasing level of importance (panel b.2).

@@

The wave-3 pattern from Figure \ref{fig:2d-regr} panel b.2 is very
similar to the teleconnection pattern associated with the ENSO. Indeed,
\citet{fogt2011} showed that there is a significant relationship between
the SAM and the ENSO. The correlation between the full SAM and the ENSO
as measured by the Multivariate Enso Index \citep{wolter2011} is -0.19.
This relationship is captured entirely the Asymmetric SAM, as this index
has a partial correlation of -0.27 with the MEI, whereas the Symmetric
SAM has null partial correlation with the MEI.

@@

\begin{figure*}
\includegraphics{regr-air-season-1} \caption[Regression pattern of surface temperature with Asymmetric and Symmetric SAM]{Regression pattern of surface temperature with Asymmetric and Symmetric SAM. P-values smaller than 0.05 (controlling for Flase Detection Rate) as hatched areas. Gray areas have more than 15\% of missing data.}\label{fig:regr-air-season}
\end{figure*}

@@

Regression of the SAM indicies with seasonal mean precipitation are
shown Figures \ref{fig:pp-regr-oceania} and \ref{fig:pp-regr-america}
for Australia and New Zealand, and South America respectively. (We
didn't detect any significant signal in South Africa.)

\begin{figure*}
\includegraphics{pp-regr-oceania-1} \caption[Same but for oceania]{Same but for oceania}\label{fig:pp-regr-oceania}
\end{figure*}

In Australia (Figure \ref{fig:pp-regr-oceania}), the annual-level
regression shows that the Full SAM is associated with a statistically
significant increase in precipitation in the Southeastern region (panel
a.1), which reproduces the results from \citet{gillett2006}. The
separation between Asymmetric and Symmetric SAM suggest that this
increase is explained by the Symmetric SAM only in the East coast (panel
c.1), which is consistent with the increased easterly flow clearly seen
in relation with this index. The Asymmetric SAM appears related to
increased precipitation in the West coast of Southeastern Australia
(panel b.2), explained by the anomalous \emph{westerly} circulation
transporting moist air to the continent.

The seasonal-level regressions show statistically significant anomalies
only in SON, with a pattern similar to the annual-level regression
(panel a.5). Panels b.5 and c.5 don't show a clear separation between
the Asymmetric and Symmetric SAM. If anything, the positive and more
significant regression coefficients in panel b.5 vs pane c.5 would
suggest more influence of the Asymmetric than the Symmetric SAM, going
against the interpretation gathered from the annual-level regressions.
This Spring signal is broadly consistent with \citet{hendon2007}, but
whereas \citet{hendon2007} also detected a strong signal in Summer,
panel a.2 shows no statistically significant association (although the
coeffcients have the consistent sign).

\begin{figure*}
\includegraphics{pp-regr-america-1} \caption[Same but for america]{Same but for america}\label{fig:pp-regr-america}
\end{figure*}

In South America (Figure \ref{fig:pp-regr-america}) , the annual-level
regression shows that the SAM is associated with statistically
significant precipitation decreases both in Southeastern South America
and Southern Chile (panel a.1). Both signals were observed by
\citet{silvestri2009} and the former also by \citet{gillett2006}. Panels
b.1 and c.1 show a remarkable clean separation between the Asymmeric SAM
--associated with the signal located in Southestern South America-- and
the Symmetric SAM --associated with the signal in Chile. This separation
is consistent with the mechanisms responsible for these effects. In
Southeastern South America, anomalous meridional winds lead to less
precipitation by inhibiting moisture convergence from the South American
Low Level Jet \citep{silvestri2009}. In Southern Chile, the reduced
westerly flow reduce moisture transport from the Pacific Ocean (cita??).
The is a small area of increased precipitation with SAM near central
Argentina which is also present in the station-based analysis by
\citet{gillett2006} and that is explained by the Asymmmetric SAM.

The negative relationship between precipiation and SAM index in
Southeastern South America and Southern Chile is evident in all seasons
except Winter (although not always stastitically significant) and it
appears to maximise in Autumn.

\subsubsection{Sea ice}

\begin{figure*}
\includegraphics{regr-ice-1} \caption[Seasonal regression of SAM indices with sea ice concentration]{Seasonal regression of SAM indices with sea ice concentration. \#FIXME}\label{fig:regr-ice}
\end{figure*}

Regressions between the Full SAM index and Antarctic Sea Ice
Concentrations (Figure \ref{fig:regr-ice}) show a great deal of
variability across seasons. The only statistically significant signal is
in Spring, when we observe negative concentraion anomalies in the
Northen Weddell Sea (panel a.4) explained by the Asymmetric SAM (panel
b.4). Both in Winter and in Spring the Asymmetric SAM is associated with
bigger Sea Ice Concentration anomalies in West Antarctica than East
Antarctica, with generally decreased concentration East of the Antarctic
Peninsula and increased concentration to the West, as expected from the
anomalous circulation correlated with this index. The Symmetric SAM
signal appears more evenly distributed accross the whole ice sheet.

\subsection{Conclusions}

\citet{silvestri2009} showed that impacts linked to the SAM changed
rather dramatically before and after 1980, and that this change can be
explained by changes in the regional-scale structure of the SAM-related
circulation. In particular, the decreased in precipitation in South
America (Figure \ref{fig:pp-regr-america} panel a.1) was absent or even
switched sign.

\citet{silvestri2009} also thowed

Therefore, is very likely that if we were to repeat this analysis using
pre-satellite data, the resulting Asymmetric SAM would look very
differnt.

@@

\citet{hosking2013} observed a pattern similar to the one near the
Amundsen Sea in panel a.4 associated with the strengthening of the ASL.
The positive anomalies of SIC near the Amundsen sea being inconsistent
with the anomalies of meridional wind associated with the ASL. Row 4 in
Figure \ref{fig:regr-ice} might shed some light to resolve this mistery.
The SAM is evidently related to the ASL through it's Asymmetric
component, and since the ASymmetric SAM is correlated with the Symmetric
SAM, the ALS will be rleated to the symemtric SAM. Consiering the
similarities between the the SIC anomalies observed by
\citet{hosking2013} (in its Figure 7) and Figure \ref{fig:regr-ice}
panel a.4, it is very likely that what was observed in that paper was a
residual effect of the SAM. \#FIXME!!
